Algebraic Groups\vspace{-1ex}

### 2019-10-23 Algebraic Geometry/Algebraic-Geometry.tex
update

>>>>

\section{October 23rd, 2019}
The plan moving forward: we are going to talk about some more ideas not from the book about topology and other topics.
Then we will rejoin the book with chapter three (skipping the discussion of dimension theory).

Here's where we're aiming: last time we talked about what it meant for a Noetherian topological space (or variety) to be irreducible.
In particular for any Noetherian topological space.
\begin{thm}\label{thm:decomp}
	$X$ admits a decomposition 
	\[X=X_1\cup\cdots\cup X_n\]
	with each $X_i$ irreducible and for all $i\ne j$, $X_i\not\subseteq X_j$.

	Furthermore, given two such decompositions 
	\[X=X_1\cup\cdots\cup X_n=Y_1\cup\cdots\cup Y_m\]
	then there exists a function $\iota:[n]\to[m]$ such that $X_i=Y_{\iota(i)}$.
\end{thm}
Recall that when $X=\Spec A$, the irreducible components $X_i$ of $X$ correspond precisely to the minimal primes $\frakp\lhd A$.

\subsection{Noetherian Induction}
Let $X$ be a Noetherian topological space and let $P$ be a set of closed subsets of $X$. Suppose that for all $Y\subseteq X$ closed, if 
\[\text{for all closed $Z\subsetneq Y$, $Z\in P$ implies that $Y\in P$ }\]
Then $X\in P$
\begin{prf}
	By contrapositive: assume that $X\in P^c.$ But now we can use axiom of choice: there exists a minimal $W\in P^c$. 
	But then for any closed $Z\subsetneq P$, we have $Z\in P$, so $W\in P$, a contradiction. But then $P^c=\varnothing=X$.
\end{prf}

Now we can prove theorem~\ref{thm:decomp}
\begin{prf}
	$\varnothing$ has the empty decomposition. Thus $\varnothing\in P$. Now if $Y\subset X$ is closed, then either $Y$ is irreducible and $Y\in P$ 
	(``admits a decomposition'') or else $Y=Z\cup W$ and assuming each are in $P$ we get that $Y\in P$ by concatenating (and eliminating). Thus by Noetherian induction we get our decomposition.

	To get the last bit, take two decompositions. Then consider 
	\[X_1=X_1\cap X=(Y_1\cap X_1)\cup\cdots\cup(Y_m\cap X_1).\]
	Then since $X_1$ is irreducible, $X_1\subseteq X_1\cap Y_j$ for some $j$ and thus $X_1\subseteq Y_j$. But then similarly $Y_j\subseteq X_i$ for some $i$.
	But by assumption on the $X_k$, $X_1=Y_j=X_i$, giving us our $\iota$.
\end{prf}

\subsection{Important Morphisms}
\begin{defn}
	A morphism $f:X\to Y$ of varieties is 
	\begin{itemize}
		\item \textbf{affine} if there exists an affine covering $U_i\subseteq Y$ such that $f^{-1}(U_i)\subseteq X$ is affine.
		\item \textbf{finite} if there exists an affine covering $U_i\subseteq Y$ such that $f^{-1}(U_i)$ is affine and 
		\[\O_Y(U_i)\to\O_X(f^{-1}(U_i))\]
		is a finite ring extension.
	\end{itemize}
\end{defn}
\begin{prop}
	$f:X\to Y$ is affine (resp. finite) if and only if for all $U\subseteq Y$ affine, $f^{-1}(U)$ is affine (resp. $f^{-1}(U)$ is affine 
	and $\O_X(U)\to \O_X(f^{-1}(U))$ is finite).
\end{prop}

### 2019-10-23 Algebraic Groups/Algebraic-Groups.tex
update

>>>>

\section{October 23rd, 2019}
Recall the following definition/proposition:
\begin{defn}
	Let $S=\Spec R$ be Noetherian. Let $G\to S$ be an affine group scheme that is flat and of finite type. Let $G$ be a scheme over $S$ with a right $G$-action.

	Then the following are equiavlent:
	\begin{itemize}
		\item $P\to S$ is a \textbf{$G$-torsor}
		\item $P\to S$ is faithfully flat and of finite type and $P\times_SG\xrightarrow{(\sigma,\pi_1}P\times_S P$ is an isomorphism.
		\item There exists $S'\twoheadrightarrow S$ faithfully flat such that 
		\begin{center}
			\begin{tikzcd}
				G\times_S S'\cong P\times_S S'\ar[r]\ar[d] & P\ar[d]\\
				S'\ar[r] & S
			\end{tikzcd}
		\end{center}
		commutes (where the isomorphism shown is $G\times_SS'$-equivariant.)
	\end{itemize}
\end{defn}

\subsection{Some Examples}
We have the trival torsor $P=G\to S$. It is a propositoon that $P\to S$ is trivial iff there exists a section $s:S\to P$.

Let $L/K$ be a finite Galois extension. Then we get $\operatorname{Gal}(L/K)$ acting on $P=\Spec L\to \Spec K$ is a $G=\operatorname{Gal}(L/K)$-torsor.
Then in the diagram in the definition above, $\Spec L$ plays the part of $S'$. Then we get that $L\otimes L\cong L[x]/f\cong \prod_{g\in G}L$ and 
\[G\times_{\Spec K}\Spec L\cong \Spec(L\otimes L)\cong \sqcup_{g\in G}\Spec L.\]

Now let $X$ be a scheme and let $\Gm$ act on a line bundle $L\to X$ with section $o:X\to L$. Then $(L\setminus o(X))\to X$ is a $\Gm$ torsor.
It is a result, although we don't have the machinery yet, that 
\begin{prop}
	There is a bijection between line bundles on $X$ and $\Gm$-torsors.
\end{prop}
\begin{rmk}
	We can do womething similar with any vector bundle over $X$: if $V\to X$ is one, then $V_x$ over $x\in X$ is a vector space. 
	We just send $V$ to $\operatorname{Frame}(V)$, which over any $x\in X$ we have the set of orderd bases of $V_x$. This gives us a $\GL_n$-torsor.
\end{rmk}

If we have a subgroup (say of an algebraic group) $H\subseteq G$, recall that we wanted to show the existence of an $H$-torsor $G\to G/H$.

We begin by talking about \textit{abstract groups}. Assume we have an exact sequence 
\[1\to K\to G\xrightarrow{\pi}Q.\]
Then $G\times K\xrightarrow{\sim}G\times_QG$ via the map $(g,k)\mapsto(g, gk)$. The proof isn't too bad.

So now consider a geometric group. If we have the same exact sequence of \textit{algebraic groups over $k$}, we get $K=G\times_Q\Spec k$(??)
and then evaluating at any scheme $S$, we get 
\[1\to K(S)\to G(S)\to Q(S)\]
and consider the map $G(S)\times K(S)\to G(S)\times_{Q(S)}G(S)$ as above. Then by Yoneda we get an isomorphism $G\times K\cong G\times_Q G$ of schemes.

\begin{cor}
	If $\pi:G\to Q$ is a faithfully flat map of linear algebraic groups over $k$, then $G\to Q$ is a torsor under $K=\ker\pi$
\end{cor}
Let's do even better!
\begin{cor}
	Let $\pi:G\to Q$ be dominant (i.e. the image is dense in $Q$) and furthermore that $Q$ is reduced. Then $G\to Q$ is faithfully flat and in particular $G\to Q$ is a $K$-torsor.
\end{cor}

\begin{prf}
	(Of the second corollary): We use the idea of ``generic flatness''. That is there exists a $U\subseteq Q$ such that $\pi^{-1}(U)\to U$ is flat. Then we can translate this by 
	the $G$-action (after passing to the algebraic closure of $k$ so that the points are dense) and then flat descent gives us the result we want. :)
\end{prf}

\begin{thm}
	Let $G$ be alinearly algebraic group over $k$. Let $X$ be a scheme over $k$ of finite type.
	Then
	\[\{G\text{-bundles on $X$}\}\cong H_{fl}^1(X,G)\]
	where $H_{fl}$ is flat cohomology.
\end{thm}
The idea here is clear when $G=\Gm$: you get connections between line bundles on $X$ (i.e. the Picard group of $X$) and $\Gm$-torsors and similarly between the line bundles and $H_{Zar}^1(X,\O_X^\times)$, 
using the (usual) Zariski sheaf cohomology.
